K Nearest Neighbour Search Using Space-filling Curves

### 2013-12-07 paper/paper.tex
added the new algorithms to the paper

@@

From one point in the Hilbert curve to another since only one bit changes by onepoint to another there is a formula to determine which bit changes. The formula for this returns the bit that will change next~\ref{g}.
\begin{equation}
  \label{g}
  g(i) = \lfloor \log _{2} \left( gc(i) \oplus gc(i+1) \right) \rfloor + 1
\end{equation}

The direction of the unit Hilbert curve is represented by~\ref{d}.
\begin{equation}
  \label{d}
  d(i) = 
    \begin{cases}
      0 & \quad \text{if $i = 0$}\\
      g(i - 1) \pmod{n} & \quad \text{if $i$ is even} \\
      g(i) \pmod{n} & \quad \text{if $i$ is odd}
    \end{cases}
\end{equation}

The entry point of the unit Hilbert curve is represented by~\ref{e}.
\begin{equation}
  \label{e}
  e(i) =  
    \begin{cases}
      0 & \quad \text{if $i = 0$}\\
      gc(2 * \lfloor x-1/2 \rfloor) & \quad \text{otherwise} \\
    \end{cases}
\end{equation}

Formula for the bitwise right rotation is displayed in \ref{right-rotate}. The bitwise left rotation is similar.
\begin{equation}
  \label{right-rotate}
  a \rightturn i = \left[ a_{{n - 1 + i}\pmod{n}} \ldots a_{i\pmod{n}} \right]_{\left[ 2 \right]}, \text{ where } a = \left[ a_{n-1} \ldots a_0\right]_{\left[ 2 \right]} 
\end{equation}

In determining the Hilbert curve index the $l$ variable used in~\ref{hilbert-point-to-index} has each bit representing whether the point point is above or below the axis corresponding to each bit. And the $e$ and $d$ variables represent the rotation and reflections nessesary to map the current sub-Hilbert curve containing the point ${\bf p}$ to the normal orientation of sub-Hilbert curve. These values are recalculated until the sub-Hilbert curve is a unit Hilbert curve. There is also a variable $w$, which is calculated at every iteration that is ORed wit$h$ after $h$ is bit-shifted left one bit. Then after all the iterations $h$ is the index of the point ${\bf p}$. The algorithn runs in $O(m^2)$ time.

>>>>

\begin{algorithm}
  \caption{COMPARE(point ${\bf p}$, point ${\bf q}$)}
  \label{compare}
  \begin{algorithmic}[1]
    \State $x \leftarrow 0, dim \leftarrow 0$
    \ForAll{j=0 to d}
    \State $y \leftarrow \text{FirstMSB}(p_{[j]}, q_{[j]})$
    \If{$x < y$}
    \State $x \leftarrow y, dim \leftarrow j$
    \EndIf 
    \EndFor \\
    \Return $p_{[dim]} < q_{[dim]}$
  \end{algorithmic}
\end{algorithm}

\begin{algorithm}
  \caption{FirstMSB(double $a$, double $b$)}
  \label{first-msb}
  \begin{algorithmic}[1]
    \State $w \leftarrow x < y \text{ and } x < (x \mathrel{\&} y)$ \\
    \Return $w$
  \end{algorithmic}
\end{algorithm}

\begin{algorithm}
  \caption{KNN Graph Construction Algorithm}
  \label{knn-graph-alg}
  \begin{algorithmic}[1]
    \Require $\text{Mortan order compare operators: } <, > 
    \text{ and Hilbert order compare operators: } <, >$
    \Ensure $A_i \text{ contains } k \text{ nearest neighbors of } p_i \text{ in } P \text{.}$
    \State $P \leftarrow \text{QuickSort}(P, <)$
    \ForAll{$p_i$ in $P$}
    \State $A_i \leftarrow \text{SET-KNN}(p_i, \{p_{i-k}, \dots, p_{i+k} \})$
    \If{$\text{SHIFT}(p_i, \lceil \text{SET-RAD}(A_i) \rceil) < p_{i+k}$}
    \State $upper \leftarrow i$
    \Else
    \State $I \leftarrow 1$
    \While{$\text{SHIFT}(p_i, \lceil \text{SET-RAD}(A_i) \rceil) < p_{i+2^I}$}
    \State $++I$
    \EndWhile
    \State $upper \leftarrow \text{min}(i+2^I, n)$
    \EndIf
    \If{$\text{SHIFT}(p_i, - \lceil \text{SET-RAD}(A_i) \rceil) > p_{i-k}$}
    \State $lower \leftarrow i$
    \Else
    \State $I \leftarrow 1$
    \While{$\text{SHIFT}(p_i, - \lceil \text{SET-RAD}(A_i) \rceil)$}
    \State $++I$
    \EndWhile
    \State $lower \leftarrow \text{max}(i-2^I, 1)$
    \EndIf
    \If{$lower \not= upper$}
    \State $A_i \leftarrow \text{CSEARCH}(p_i, lower, upper)$
    \EndIf
    \EndFor \\
    \Return $A_i$
  \end{algorithmic}
\end{algorithm}

\begin{algorithm}
  \caption{CSEARCH(point $p_i$, int $lo$, int $hi$)}
  \label{csearch}
  \begin{algorithmic}[1]
    \If{$(hi - lo) < v $}
    \State $A_i \leftarrow \text{SET-KNN}(p_i, A_i \cup \{ p_{lo} \dots p_{hi} \})$
    \Return $A_i$
    \EndIf
    \State $mid \leftarrow (hi+lo)/2$
    \State $A_i \leftarrow \text{SET-KNN}(p_i, A_i \cup p_{mid})$
    \If{$\text{DIST}(p_i, \text{CIRCLE}(p_{lo}, p_{hi})) \ge \text{SET-RAD}(A_i)$} 
    \Return $A_i$
    \EndIf
    \If{$p_i < p_{mid}$}
    \State $A_i \leftarrow \text{CSEARCH}(p_i, lo, mid-1)$
    \If{$p_{mid} < \text{SHIFT}(p_i, \lceil \text{SET-RAD}(A_i) \rceil)$}
    \State $A_i \leftarrow \text{CSEARCH}(p_i, mid+1, hi)$
    \EndIf
    \Else
    \State $A_i \leftarrow \text{CSEARCH}(p_i, mid+1, hi)$
    \If{$\text{SHIFT}(p_i, - \lceil \text{SET-RAD}(A_i) \rceil) < p_{mid}$}
    \State $A_i \leftarrow \text{CSEARCH}(p_i, low, mid - 1)$
    \EndIf
    \EndIf \\
    \Return $A_i$

@@

From one point in the Hilbert curve to another since only one bit changes by onepoint to another there is a formula to determine which bit changes. The formula for this returns the bit that will change next~(\ref{g}).
\begin{equation}
  \label{g}
  g(i) = \lfloor \log _{2} \left( gc(i) \oplus gc(i+1) \right) \rfloor + 1
\end{equation}

The direction of the unit Hilbert curve is represented by~(\ref{d}).
\begin{equation}
  \label{d}
  d(i) = 
    \begin{cases}
      0 & \quad \text{if $i = 0$}\\
      g(i - 1) \pmod{n} & \quad \text{if $i$ is even} \\
      g(i) \pmod{n} & \quad \text{if $i$ is odd}
    \end{cases}
\end{equation}

The entry point of the unit Hilbert curve is represented by~(\ref{e}).
\begin{equation}
  \label{e}
  e(i) =  
    \begin{cases}
      0 & \quad \text{if $i = 0$}\\
      gc(2 * \lfloor x-1/2 \rfloor) & \quad \text{otherwise} \\
    \end{cases}
\end{equation}

Formula for the bitwise right rotation is displayed in~(\ref{right-rotate}). The bitwise left rotation is similar.
\begin{equation}
  \label{right-rotate}
  a \rightturn i = \left[ a_{{n - 1 + i}\pmod{n}} \ldots a_{i\pmod{n}} \right]_{\left[ 2 \right]}, \text{ where } a = \left[ a_{n-1} \ldots a_0\right]_{\left[ 2 \right]} 
\end{equation}

In determining the Hilbert curve index the $l$ variable used in~(\ref{hilbert-point-to-index}) has each bit representing whether the point point is above or below the axis corresponding to each bit. And the $e$ and $d$ variables represent the rotation and reflections nessesary to map the current sub-Hilbert curve containing the point ${\bf p}$ to the normal orientation of sub-Hilbert curve. These values are recalculated until the sub-Hilbert curve is a unit Hilbert curve. There is also a variable $w$, which is calculated at every iteration that is ORed wit$h$ after $h$ is bit-shifted left one bit. Then after all the iterations $h$ is the index of the point ${\bf p}$. The algorithn runs in $O(m^2)$ time.
